A Summery about reasoning with logic

### 2015-11-10 methods/agents/summery.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

\begin{tabular}{lll}
	Modal & $\Leftrightarrow$ & Interpertation \\ \toprule
$M,s \vDash p$ & $\Leftrightarrow$ & $\pi(s,p) = tt$ \\
$M,s \vDash \phi \wedge \varphi $ & $\Leftrightarrow$ & $ M,s \vDash \phi
\mbox{ and }M,s\vDash \varphi$ \\
$M,s \vDash \Box \phi $ & $\Leftrightarrow$ & $ M,t \vDash \phi $ for 
\emph{every} $t$ such that $R(s,t)$ \\
$M,s \vDash \Diamond \phi $ & $\Leftrightarrow$ & $ M,t \vDash \phi $ for 
\emph{some} $t$ such that $R(s,t)$ \\
\end{tabular}

@@

$\phi$ is \emph{valid} wrt class $\zeta$ &
$\zeta\vDash\phi$ & $\Leftrightarrow$ & $M\vDash\phi$ for all Kripke models
$M \in \zeta$ \\
\end{tabular}

@@

\section{APL}
Medhi dastani taught about the agent programming langauge (APL).
\subsection{Agent data structure}
This represents the mental state of a agent. \\
\begin{tabular}{ll}
	Name & Does what \\ \toprule
	Beliefs & Information available to the agent \\
	Goals & Objectives that the agent wants to achieve \\
	Events & Obsevations of (environmental) changes \\
	Capabilities & Actions that the agent can perform \\
	Plans & Procedures to achieve objectives \\
	Reasoning rules & Reason about goal and plans \\
\end{tabular} \\

>>>>

\begin{tabular}{lll}
	Modal & $\Leftrightarrow$ & Interpertation \\ \toprule
$M,s \vDash p$ & $\Leftrightarrow$ & $\pi(s,p) = tt$ \\
$M,s \vDash \phi \wedge \varphi $ & $\Leftrightarrow$ & $ M,s \vDash \phi
\mbox{ and }M,s\vDash \varphi$ \\
$M,s \vDash \Box \phi $ & $\Leftrightarrow$ & $ M,t \vDash \phi $ for 
\emph{every} $t$ such that $R(s,t)$ \\
$M,s \vDash \Diamond \phi $ & $\Leftrightarrow$ & $ M,t \vDash \phi $ for 
\emph{some} $t$ such that $R(s,t)$ \\\bottomrule
\end{tabular}

@@

$\phi$ is \emph{valid} wrt class $\zeta$ &
$\zeta\vDash\phi$ & $\Leftrightarrow$ & $M\vDash\phi$ for all Kripke models
$M \in \zeta$ \\\bottomrule
\end{tabular}

@@

\subparagraph{Caution} the necisitation is different from the invalid
assertion: $\phi \to \Box \phi$.

\subsubsection{Derivability in K}
A \emph{derivation} of a formula $\phi$ is a finite sequence of formulas
$\phi_1, \dots \phi_n = \phi$. where each $\phi_i$ for $1 \ge i \ge n$
is either an instance of the acioms (or rather axiom schemes), or the
conclusion of one of the rules of which the premises have
been derived already, i.e{.} appear as $\phi_j$ in the sequance with
$j < i$.

% sad thing is I understand this crap now.
When we can derive an epistemic formula $\phi$ by using the axioms
and rules of $K_{(m)}$ we write $K_{(m)} \vdash \phi$

System $K_{(m)}$ is \emph{sound and complete}. This means
that exactly ll valid modal assertions can be obtained by derivations in
system $K_{(m)}$.

\[\vDash \phi \Leftrightarrow K_{(m)} \vdash \phi\]

% I leave out the (non) theoroms, it seems overkill.

\subsection{Dynamic logic}
An example of an (indexed) version of system $K$ is dynamic logic,
wehre the $\Box$ modality is associated with the execution results of
a program / action.

$\Box_\alpha$, normally written $[\alpha]$ 
with reading $[\alpha]\phi$: after exuction of $\alpha$ it holds that $\phi$

\[<\alpha>\phi = \neg [\alpha] \neg \phi \]

\subsubsection{Semantics}
Accessibility relation $R_\alpha$ for every action $\alpha$

\begin{tabular}{llll}
	% I added the natural collumn, because otherwise I won't remember.
	% Translating something in a different notation is stupid.
	notation & = & definition & natural \\ \toprule
	$R_{\alpha;\beta}$ & = & $R_\alpha \circ R_\beta$ &
	execute from $R_\alpha$ until including $R_\beta$\\
	$R_{\alpha+\beta}$ & = & $R_\alpha \cup R_\beta$ &
	execute $R_\alpha$ or $R_\beta$\\
	$R_{\alpha*}$ & = & $R_\alpha*$ &
	Execute an arbitrary times \\ \bottomrule
\end{tabular}

\subsubsection{Interpertation}
\begin{tabular}{lll}
	Modal & $\Leftrightarrow$ & Interpertation \\ \toprule
	$M,s \vDash [\alpha]\phi$ & $\Leftrightarrow$ &
	for all $s'$ with $R_\alpha(s,s'): M,s' \vDash \phi$ \\
	$M,s \vDash <\alpha>\phi$ & $\Leftrightarrow$ &
	for some $s'$ with $R_\alpha(s,s'): M,s' \vDash \phi$ \\\bottomrule
\end{tabular}

\subsection{Special properties of accessibility relations}
\begin{tabular}{lll}
	R is \dots& if\dots & natural language \\ \toprule
	reflexive & $\forall s \in S(s,s) \in R$ & all s are linked with 
	themselves\\

	transitive & $\forall s,t,u \in S:(s,t) \in R \& (t,u) \in R \Rightarrow
	(s,u) \in R$ & \brcell{if s is related to t, and t is related to u,  \\
		then s has a relation with u, \\ in other words, you can make shortcuts}\\

	symmetrical &$\forall s,t \in S:(s,t) \in R \Rightarrow (t,s) \in R$ &
	undirected, can go back and forth\\
	euclidiean & $\forall s,t,u \in S:(s,t) \in R \& (s,u) \in R \Rightarrow
	(t,u) \in R$ & \brcell{if s has a relation with t and u, then u and t \\
	have a relation to, you can make triangles.} \\
	serial & $\forall s \in S \exists t \in S (s,t) \in R$&
	\brcell{All states in S are connected to \\ at least one other state in S}\\
	an equivalance relation& R is \emph{reflexive, transitive and 
	symetrical}& \brcell{the states R links to have something in \\ 
	common (not equal?)} \\ \bottomrule
\end{tabular}
\section{APL}
Medhi dastani taught about the agent programming langauge (APL).
\subsection{Agent data structure}
This represents the mental state of a agent. \\
\begin{tabular}{ll}
	Name & Does what \\ \toprule
	Beliefs & Information available to the agent \\
	Goals & Objectives that the agent wants to achieve \\
	Events & Obsevations of (environmental) changes \\
	Capabilities & Actions that the agent can perform \\
	Plans & Procedures to achieve objectives \\
	Reasoning rules & Reason about goal and plans \\
\end{tabular} \\

### 2015-11-10 methods/cogniveProcessing/summery.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

\section{Langauge}
Trying to figure out what langauge is.

@@

>>>>

\section{Langauge structure}
Trying to figure out what langauge is.

@@

\subsection{Felicitous}
Helpfull to the conversation. A sentence can be this or not. This
works better than a simple truth value, because it can also handle
state changing sentences, such as: ``I declare war''.

\subsection{Cooperative principle}
A conversation has a goal, that the speaker wants to contribute to, assuming
the hearer does the same.

\subsubsection{Quality}
Speak the truth, and don't say things you don't have evidence for.

\paragraph{Example} B has stolen A's chocoletes
A: ``have you stolen my chocolates?''\\
B1: {\color{green} ``Yes.''}\\
B2: {\color{red} ``No.''}

\subsubsection{Quantity}
Don't give to much or too little information

\paragraph{Example} B knows A's chocolates are on the table.
A: ``Where are my chocolates?''\\
B1: {\color{green} ``On the table.''}\\
B2: {\color{red} ``Somewhere.''}\\
B2: {\color{red} ``On the table. Which is an oaken item of furniture
located on the other end of this living room''}

\subsubsection{Relevance}
Make relevant contributions.

\paragraph{Example} B has not stolen A's chocolates.
A: ``Someone has stolen my chocolates?''\\
B1: {\color{green} ``That's terrible.''}\\
B2: {\color{red} ``The weather's really nice today, don't you agree?.''}\\

\subsubsection{Manner}
Be clear and bried, don't do obfuscation.

\paragraph{Example} 
A: ``Tell me about the performance''\\
B1: {\color{green} ``Mrs Smith sang \emph{Bridge over Troubled Water.}''}\\
B2: {\color{red} ``Mrs Smith produced a series of soundwaves that resembled
the melody of \emph{Bridge over Troubled Water}.''}\\

\subsection{Impliciture}
A: ``Where cna I find a service station?''\\
$B_1$: ``Around the corner. \emph{My mother wokrs there''}.\\
$B_2$: ``Around the corner. \emph{But I'm not sure wether it's open''}.

Both $B_1$ and $B_2$ violate quantity, but unlike $B_1$, $B_2$ is 
felicitous. Why?

If B had known the service station would be closed and useless
to A his remark would have violated relavance.

So A concludes: ``B thinks that the service station is open''.

B, knowing this preemts this conclusion by adding ``but I'm not sure it's 
open'', so violating a maxim cna be communicatively meaningfull (to preemt
wrong conclusions).

So B thinks about how his message will be received by A, and decides that
it will be interpeted wrongly so he adds information.

An inference like $A's$ (``$B$ thinks the service station is open'')
is called an \emph{implicature}

\subsubsection{Two flavours}
There are two flavours of implicuture: \emph{conevnetional} and 
conversational.

\paragraph{Conversational} Implicatures: the information that conversational
partners infer from each other's utterances, based on the gricean maxims.

\subparagraph{Not entailment!} Conversational implicatures can be canceled.

\paragraph{Conventional} implicatures: non-cancellable implicutres
that are part of the conceptual meaning of a word.

### 2015-11-10 methods/reasoning/huiswerk/inl1026.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

\begin{tikzpicture}
[scale=.8,auto=left,every node/.style={circle,fill=blue!20}]
\draw (3,3) circle (3.5);
\node (n1) at (1,3) {$A \Rightarrow A$};
\node (n2) at (3,5) {$B \Rightarrow A$};
\node (n3) at (4,2)  {$C \wedge B, A \Rightarrow A$};
\foreach \from/\to in {n1/n2,n2/n3}
\draw (\from) -- (\to);

\end{tikzpicture}

, if
 If $C$ $I(\Gamma \Rightarrow A)=1$ then
so does $I(\Pi, \Box \Gamma \Rightarrow \Box A, \Delta) = 1$ for example:
\begin{tikzpicture}
[scale=.8,auto=left,every node/.style={circle,fill=blue!20}]
\draw (3,3) circle (3.5);
\node (n1) at (1,3) {$A \Rightarrow A$};
\node (n2) at (3,5) {$B, A \Rightarrow A$};
\node (n3) at (4,2)  {$C \wedge B, A \Rightarrow A$};
\foreach \from/\to in {n1/n2,n2/n3}
\draw (\from) -- (\to);

\end{tikzpicture}

@@

As an intial observation the rules of $GKC_{K4}$ and $G$ are functionally
the same. They're both intendid to remove the box on the right side and to
hide it away in the left side in one of the `garbage' spots $(\Pi$ and 
$\Delta)$ of the axioms.

However the $GKC_{K4}$ has also garbage spots in the conclusion. So to show
something is proofable in $GKC_{K4}$ and not in $G$ all you need to do is
fill in one of those garbage spots. Consider the following sequant:

\[A, \Box B \Rightarrow \Box B, A\]

@@

\begin{prooftree}
	\AxiomC{}
	\RightLabel{Ax}
	\UnaryInfC{$\Box B, B \Rightarrow B$}
	\RightLabel{$R$}
	\UnaryInfC{$A, \Box B \Rightarrow \Box B, A$}
\end{prooftree}

>>>>

\begin{tikzpicture}
[scale=.8,auto=left,every node/.style={circle,fill=blue!20}]
\draw (3,3) circle (3.5);
\node (n1) at (1,3) {$A$};
\node (n2) at (3,5) {$B$};
\node (n3) at (4,2)  {$C \wedge B$};
\foreach \from/\to in {n1/n2,n2/n3}
\draw (\from) -- (\to);
\end{tikzpicture}

In this model you can see that the preceding sequant does not automaticly
imply an A. So the premise ($\Gamma \Rightarrow A$) is false, 
which means the conclusion is also false,
there is no path available form every state to an A.

 If $C$ $I(\Gamma \Rightarrow A)=1$ then
so does $I(\Pi, \Box \Gamma \Rightarrow \Box A, \Delta) = 1$ for example:
\\
\begin{tikzpicture}
[scale=.8,auto=left,every node/.style={circle,fill=blue!20}]
\draw (3,3) circle (3.5);
\node (n1) at (1,3) {$A$};
\node (n2) at (3,5) {$A \wedge B$};
\node (n3) at (4,2)  {$A \wedge (C \vee B)$};
\foreach \from/\to in {n1/n2,n2/n3}
\draw (\from) -- (\to);
\end{tikzpicture}
\\
In this model every state A holds at least.
So there is also always a path available to a state were A holds.
This means the premise ($\Gamma \Rightarrow A$) and conclusion
($\Pi, \Box \Gamma \Rightarrow \Box A, \Delta$) are true.

@@

As an intial observation the rules of $GKC_{K4}$ and $G$ are functionally
the same. They're both intendid to remove the box on the right side and to
hide it away in the left side in one of the `garbage' spot ($\Gamma$) of
the axioms.

However the $GKC_{K4}$ has also garbage spots $(\Pi$ and
$\Delta)$ in the conclusion. So to show
something is proofable in $GKC_{K4}$ and not in $G$ all you need to do is
fill in one of those garbage spots. Consider the following sequant:

\[A, \Box B \Rightarrow \Box B\]

@@

\begin{prooftree}
	\AxiomC{}
	\RightLabel{Ax}
	\UnaryInfC{$\Box B, B \Rightarrow B$}
	\RightLabel{$R$}
	\UnaryInfC{$A, \Box B \Rightarrow \Box B$}
\end{prooftree}

### 2015-11-10 methods/reasoning/summery.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

\subsection{Head of a proof}
In the head of the proof you don't do any work, you just state what we're
talking about and what you're going to do. So you start with the definitions.
Then you say what you're going to proof (by using the defintions), so here
you just repeat the question. And finally you mention your strategy.

@@

@@

>>>>

\subsection{Head of a proof}
In the head of the proof you don't do any work, you just state what we're
talking about and what you're going to do. So you start with the definitions.
Remeber the definitions should be stupid, only add obvious things, don't do
deductions until the body of the proof.
Then you say what you're going to proof (by using the defintions), so here
you just repeat the question. And finally you mention your strategy.

@@

So for the base case the tabloo is method is used to show that the valuation
trinceling down happens for the rules, then induction is used for the entire
proof tree to show that it works for the entire system.

@@

\paragraph{Proof that for $L\to$ of GKC that if one of the premises is not
true (under all valuations) than so is the conclusion}

\subparagraph{Definitions} GKC is gentzen calculy, in which $L\to$ is defined
as follows:

\begin{prooftree}
	\AxiomC{$\Gamma \Rightarrow  A, \Delta$}
	\AxiomC{$\Gamma, B \Rightarrow \Delta$}
	\RightLabel{$ L\to$ }
	\BinaryInfC{$\Gamma, A \to B \Rightarrow \Delta$}
\end{prooftree}

Where $\Gamma$ is a collection of formulas connected by $\wedge$ and $\Delta$
is a collection of formulas connected by $\vee$.
A and B are singular formulas.
I(x) is a function that changes the $,$ into proper connectives. So in case
of the antecedent it will become a $\wedge$ and in the succedent a $\vee$.

\subparagraph{Strategy} Under all valuations, if one of the premises is not
true of $L\to$ then so is the conclusion. This will be proven with help
of the tabloo method.
\subparagraph{Proof}
There are two cases to consider first $V(I(\Gamma \Rightarrow  A, \Delta))=0$
and secondly $V(I(\Gamma, B \Rightarrow  \Delta))=0$

$V(I(\Gamma \Rightarrow  A, \Delta))=0$ implies $V(A)=0$ and for all formulas
where $C\in \Delta$ and $D \in \Gamma$, $V(C) = 0$ and $V(D)=1$. In this case
$V(A \to B) = 1$ and thus $V(I(\Gamma, A \to B \Rightarrow \Delta))=0$.

\emph{So here I say if the $A$ premise is false then the conclusion is false
	because propositional logic, you need the $I$ function to deal with
	commas. And you need the V function to execute a true false valuation
(ergo execute propositional logic)}

$V(I(\Gamma, B \Rightarrow \Delta))=0$ implies $V(B)=1$ and for all formulas
where $C\in \Delta$ and $D \in \Gamma$, $V(C) = 0$ and $V(D)=1$. In this case
$V(A \to B) = 1$ and thus $V(I(\Gamma, A \to B \Rightarrow \Delta))=0$.

\emph{So step by step, say what the valuation for the premis implies for the
	individual elements (if its a set of elements, just do it on a member of
the set, then show it for the conclusion with help of propositional logic}
